% ===== Appendix Z (file: src/appendices/Z-archaios.tex) =====
\section{ARCHAIOS scoring note (public flavor)}
\label{app:archaios}

This appendix briefly documents the public (non-proprietary) metrics
reported by the ARCHAIOS pipeline for the current version of the note.
The public flavor exposes \emph{metric values only}
(with a short rationale); private recommendations are intentionally omitted.

\subsection*{Z.1. Reported coefficients}
The following six coefficients are computed from the LaTeX/PDF artifact
using structural and textual cues:
\begin{itemize}
  \item \textbf{KPD} — Proof plausibility;
  \item \textbf{KGD} — Logical coherence;
  \item \textbf{KSD} — Structural depth;
  \item \textbf{KSR} — Source robustness;
  \item \textbf{KFZ} — Formal compliance;
  \item \textbf{KKA} — Clarity \& accessibility.
\end{itemize}
A composite confidence score accompanies the metrics.

\subsection*{Z.2. Where the numbers come from}
The exact values are saved by CI to the ancillary folder as JSON and CSV:
\begin{center}
  \texttt{ancillary/archaios\_public\_report.json},\quad
  \texttt{ancillary/archaios\_public\_report.csv}.
\end{center}
These files are generated from the rendered public report (\texttt{public.md})
in the private evaluation job and copied to this repository during the
\texttt{archaios-publish} workflow.

\subsection*{Z.3. Minimal preview table}
For convenience we include a one-row preview (rounded to one decimal).
Values below are illustrative; the canonical source of truth is the
CSV/JSON files mentioned above.
\begin{center}
\begin{tabular}{l|r}
\hline
Metric & Score (\%) \\
\hline
KPD &  \,– \\
KGD &  \,– \\
KSD &  \,– \\
KSR &  \,– \\
KFZ &  \,– \\
KKA &  \,– \\
\hline
\end{tabular}
\end{center}

\subsection*{Z.4. Notes}
(1) Public flavor never emits prescriptive edits.  
(2) Metric definitions are stable across versions; the extraction
heuristics may evolve, which is reflected in the ancillary version history.
