% =====================================================================
% File: 02-preliminaries.tex
% Block 1 of 4 — Introduction, Goals, Invariants, Organization
% =====================================================================

\chapter{Preliminaries}

% ---------------------------------------------------------------------
% Chapter orientation
% ---------------------------------------------------------------------
\section*{Orientation and Intent}

This chapter establishes the analytic and geometric setting for the entire
monograph. Its role is foundational: it fixes once and for all the
normalizations, conventions, and persistent invariants that will be used
throughout. Every later construction (kernels, projectors, trace formulae)
will cite this chapter to avoid repetition.

The goals are fourfold:
\begin{enumerate}
  \item To specify the geometric background of hyperbolic surfaces of finite area,
        including cusps, elliptic points, and geodesic structure.
  \item To fix the sign and parameterization of the Laplace--Beltrami operator,
        together with the spectral decomposition of $L^{2}(M)$.
  \item To describe truncation conventions and cusp neighborhoods with explicit
        dependence on cusp widths and scaling matrices.
  \item To declare the Selberg/Harish–Chandra transform conventions,
        Sobolev norms, and dependency rules for error terms and constants.
\end{enumerate}

The intent is not only to provide clarity but also reproducibility:
all constants will be annotated explicitly, and dependencies will be
traceable to invariants of $(M,\Gamma)$ rather than hidden parameters.

% ---------------------------------------------------------------------
\section*{Chapter Goals}

\begin{itemize}[leftmargin=2em]
  \item[\textbf{(G1)}] Specify the hyperbolic geometry:
  the upper half-plane $\mathbb{H}$ with metric
  \[
    ds^2=\frac{dx^2+dy^2}{y^2}, \qquad
    d\mu(z)=\frac{dx\,dy}{y^2},
  \]
  and the quotient $M=\Gamma\backslash\mathbb{H}$ by a cofinite
  Fuchsian group $\Gamma\subset\mathrm{PSL}_2(\mathbb{R})$.

  \item[\textbf{(G2)}] Fix the Laplacian:
  $\Delta \ge 0$ with spectral parameter $\lambda=\tfrac14+r^2$,
  $r\in[0,\infty)$.
  This convention aligns with Selberg~\cite{Selberg1956},
  Hejhal~\cite{Hejhal1983}, and Iwaniec~\cite{Iwaniec2002}.

  \item[\textbf{(G3)}] Define cusp neighborhoods and truncations $M(Y)$,
  with canonical coordinates from scaling matrices
  and explicit relation of cusp widths $w_\mathfrak{a}$ to volume and boundary length.

  \item[\textbf{(G4)}] Record the spectral decomposition on $L^{2}(M)$:
  discrete eigenfunctions $\{\varphi_j\}$ and Eisenstein series
  $E_{\mathfrak{a}}(z,1/2+ir)$, with Plancherel measure $dr/(4\pi)$.

  \item[\textbf{(G5)}] Normalize the Selberg/Harish–Chandra transform
  for radial kernels, consistent with later kernel constructions.

  \item[\textbf{(G6)}] State uniform Sobolev bounds and mapping
  properties required for error analysis and stationary phase.

  \item[\textbf{(G7)}] Establish explicit dependency rules for constants:
  all implied constants are annotated as $O_{\Gamma,\beta}(\cdot)$,
  where $\beta$ is a lower bound for the spectral gap.
\end{itemize}

% ---------------------------------------------------------------------
\section*{Invariants and Dependency Ledger}

\begin{itemize}[leftmargin=2em]
  \item[\textbf{(I1)}] \textit{Surface and volume.}
  $M=\Gamma\backslash\mathbb{H}$ with $\vol(M)\in(0,\infty)$.

  \item[\textbf{(I2)}] \textit{Cuspidal data.}
  Number of cusps $h$, cusp widths $(w_1,\dots,w_h)$,
  and scaling matrices $\sigma_\mathfrak{a}$.
  All cusp-dependent constants are functions of this tuple.

  \item[\textbf{(I3)}] \textit{Spectral gap.}
  $\beta\in(0,\tfrac14]$ denotes a fixed lower bound for
  $r_j^2$ of nontrivial spectrum.
  Constants depending on $\beta$ are written $O_{\Gamma,\beta}(\cdot)$.

  \item[\textbf{(I4)}] \textit{Normalization of Eisenstein series.}
  Continuous spectrum integrated with measure $dr/(4\pi)$.
  Constant terms follow standard scattering conventions.

  \item[\textbf{(I5)}] \textit{Asymptotic notation.}
  $A\lesssim B$ means $A\le C\,B$ for some absolute $C$;
  $A\asymp B$ means both $A\lesssim B$ and $B\lesssim A$.
  Dependence on parameters is annotated:
  $O_X(\cdot)$ indicates dependence on $X$ only.

  \item[\textbf{(I6)}] \textit{Propagation scale.}
  The natural comparison time is $T\asymp\log\lambda$.
  Constants are fixed globally (Chapter~5) and will not affect error orders.
\end{itemize}

% ---------------------------------------------------------------------
\section*{Organization of the Chapter}

The chapter consists of modular blocks, each self-contained and audited:

\begin{itemize}[leftmargin=2em]
  \item \textbf{Geometry:} hyperbolic metric, distance, injectivity, and
        global volume conventions.
        (Input: \texttt{02b-geometry.tex})

  \item \textbf{Cusps and truncation:} cusp neighborhoods, truncation
        domains $M(Y)$, smoothed truncations, Eisenstein series, and
        Maass–Selberg relations.
        (Input: \texttt{02b-cusps.tex})

  \item \textbf{Selberg/Harish–Chandra transform:} definition, inversion,
        admissible test functions, Paley–Wiener duality, stationary phase.
        (Input: \texttt{02b-selberg-transform.tex})
\end{itemize}

Each block closes with an internal consistency check and a forward link
to later chapters. The chapter itself concludes with a global audit
summarizing achieved goals, preserved invariants, and dependency rules.

% =====================================================================
% End of File: 02-preliminaries.tex (Block 1 of 4)
% =====================================================================

% ---------------------------------------------------------------------
% Invariants and dependency ledger
% ---------------------------------------------------------------------
\subsection*{Invariants and Dependency Ledger}

In order to ensure clarity and reproducibility throughout the monograph, 
we fix here the list of persistent invariants and the precise rules by 
which constants and error terms may depend on geometric and spectral data.  
This ledger serves as a reference point for all subsequent chapters.  
Every constant introduced later must be justified in terms of the entries below.

\begin{itemize}[leftmargin=2em]

  \item[\textbf{(I1)}] \textit{Surface and volume.}  
  The basic object is the finite–area hyperbolic surface
  \[
    M = \Gamma \backslash \mathbb{H},
  \]
  where $\Gamma \subset PSL_{2}(\mathbb{R})$ is a cofinite Fuchsian group.  
  Its hyperbolic volume
  \[
    \vol(M) = \int_{\Gamma \backslash \mathbb{H}} \frac{dx\,dy}{y^{2}}
  \]
  is finite and positive, and will appear repeatedly in main terms of Weyl–type laws
  (cf. Selberg~\cite{Selberg1956}, Buser~\cite{Buser1992}).

  \item[\textbf{(I2)}] \textit{Cuspidal data.}  
  The number of cusps is $h \geq 0$.  
  To each cusp $\mathfrak{a}$ we attach a width $w_{\mathfrak{a}}$ and a scaling matrix 
  $\sigma_{\mathfrak{a}} \in PSL_{2}(\mathbb{R})$.  
  All cusp–dependent constants in this monograph are functions of the finite tuple
  \[
    (w_{1}, \dots, w_{h}; \, \sigma_{1}, \dots, \sigma_{h}).
  \]
  This convention guarantees that no hidden dependence on $\lambda$ or on truncation
  parameters is introduced.

  \item[\textbf{(I3)}] \textit{Spectral gap.}  
  Let $\{ \lambda_{j} = \tfrac14 + r_{j}^{2} \}$ denote the discrete spectrum of $\Delta$,
  excluding the trivial eigenvalue $\lambda_{0}=0$.  
  We fix once and for all a lower bound
  \[
    \beta \in (0, \tfrac14]
  \]
  such that $r_{j}^{2} \geq \beta$ for all nontrivial $j$.  
  Constants depending on $\beta$ are explicitly annotated as $O_{\Gamma,\beta}(\cdot)$.  
  This convention follows Iwaniec~\cite{Iwaniec2002} and Sarnak~\cite{Sarnak2004}.

  \item[\textbf{(I4)}] \textit{Normalization of Eisenstein series.}  
  For each cusp $\mathfrak{a}$ we adopt the convention
  \[
    E_{\mathfrak{a}}(z, s) = \sum_{\gamma \in \Gamma_{\mathfrak{a}} \backslash \Gamma}
    \Im(\sigma_{\mathfrak{a}}^{-1}\gamma z)^{s}, \qquad \Re(s) > 1,
  \]
  analytically continued to $\Re(s) \geq \tfrac12$.  
  The Plancherel measure for the continuous spectrum is fixed as $dr/(4\pi)$, and
  constant terms use the scattering matrix normalization of Hejhal~\cite{Hejhal1983}.  

  \item[\textbf{(I5)}] \textit{Asymptotic notation.}  
  We adhere to the following conventions:  
  $A \lesssim B$ means there exists an absolute $C>0$ such that $A \leq C\,B$;  
  $A \asymp B$ means both $A \lesssim B$ and $B \lesssim A$;  
  $O_{X}(\cdot)$ indicates dependence only on the parameter set $X$.  
  Whenever constants depend on $(\Gamma,\beta)$ this is made explicit via
  $O_{\Gamma,\beta}(\cdot)$.  
  No error term is allowed to hide a dependence outside this ledger.

  \item[\textbf{(I6)}] \textit{Propagation scale.}  
  In all microlocal arguments, the natural propagation time is
  \[
    T \asymp \log \lambda,
  \]
  reflecting exponential volume growth of hyperbolic balls.  
  The proportionality constants involved in this definition are fixed in Chapter~5
  and remain universal.  
  They affect no asymptotic bounds but must be recorded to ensure reproducibility.

\end{itemize}

\medskip

\noindent
\textbf{Audit of invariants.}  
The ledger above has been checked for consistency with the global glossary
(Chapter~0).  
It isolates all sources of hidden constants and ensures transparency in later
error bounds.  
Forward links: invariants (I1–I6) will be invoked in Chapters~3–6 for kernel
constructions, in Chapter~7 for error terms in theorems, and in Chapter~8
for explicit arithmetic applications.

% ---------------------------------------------------------------------
% Organization of the chapter (reader-facing map)
% ---------------------------------------------------------------------
\subsection*{Organization of the Chapter}

This chapter consists of three self-contained building blocks.  
Each block develops a distinct analytic or geometric component that will be used 
as input for the localized trace formula.  
The organization is deliberately modular: each block closes with an internal consistency check, 
and the chapter itself concludes with a full audit.  
This ensures that every convention fixed here can be cited later without repetition.

\begin{itemize}[leftmargin=2em]
  \item \textbf{Geometry of the hyperbolic plane and quotient surfaces.}  
  We recall the metric, distance function, geodesic structure, 
  volume growth, injectivity radius, fundamental domains,  
  and analytic kernels such as the heat, wave, and resolvent kernels.  
  This material provides the global background for microlocal analysis 
  and orbital decompositions.  
  (Technical input: \texttt{02b-geometry.tex})

  \item \textbf{Cuspidal structure and height truncation.}  
  We develop canonical cusp neighborhoods via scaling matrices, 
  define truncation domains $M(Y)$, compute explicit volume and boundary formulas, 
  and introduce both sharp and smoothed truncation operators $\Lambda^{Y}$ and $\Lambda^{Y}_{\mathrm{sm}}$.  
  Fourier expansions of Eisenstein series, Maass–Selberg relations, 
  scattering matrices, and effective Parseval identities are also recalled.  
  This block handles the analytic difficulties introduced by noncompactness.  
  (Technical input: \texttt{02b-cusps.tex})

  \item \textbf{Selberg/Harish–Chandra transform for radial kernels.}  
  We define the transform $q(r) \mapsto h(t)$ and its inversion, 
  record normalization of spherical functions,  
  state Plancherel and Paley–Wiener theorems, and develop examples 
  (heat kernel, wave kernel, cutoff projectors).  
  We also explain the role of admissible test functions, 
  convolution properties, orbital integrals, and stationary phase analysis.  
  This block is the analytic bridge between geometry and spectrum.  
  (Technical input: \texttt{02b-selberg-transform.tex})
\end{itemize}

\medskip

\noindent
\textbf{Pedagogical note.}  
The reader can treat this chapter as a reference manual.  
Every time a constant, a volume estimate, or a transform normalization appears in later proofs, 
its justification is located here.  
Forward links to later chapters are indicated in each block, 
and backward consistency with the global glossary (Chapter~0) is verified in the final audit.

\medskip

% ---------------------------------------------------------------------
% Inputs: technical blocks for Chapter 2
% ---------------------------------------------------------------------

% Geometry block: metric, distance, volume, injectivity considerations.
\input{sections/02-preliminaries/blocks/02b-geometry.tex}

% Cuspidal analysis: coordinates, truncation, scattering data conventions.
\input{sections/02-preliminaries/blocks/02b-cusps.tex}

% Selberg/Harish–Chandra transform: test functions and radial kernels.
\input{sections/02-preliminaries/blocks/02b-selberg-transform.tex}

% ---------------------------------------------------------------------
% Chapter Audit
% ---------------------------------------------------------------------
\subsection*{Chapter Audit}

\noindent\textbf{Scope and coverage.}  
This chapter has fixed the full analytic and geometric setting.  
The model $(\mathbb{H},ds^{2})$ and the quotient surface $M=\Gamma\backslash\mathbb{H}$ 
with finite area have been established.  
Cuspidal neighborhoods and truncations $M(Y)$ are defined with explicit formulas 
for volume and boundary length.  
The Laplacian sign convention $\Delta\ge 0$ and its parametrization $\lambda=\tfrac14+r^{2}$ 
have been fixed globally.

\medskip
\noindent\textbf{Spectral conventions.}  
We recorded the decomposition of $L^{2}(M)$ into discrete eigenfunctions 
$\{\varphi_{j}\}$ and Eisenstein series $E_{\mathfrak a}(z,\tfrac12+ir)$,  
normalized with Plancherel measure $dr/(4\pi)$.  
This convention is consistent with classical sources (Selberg~\cite{Selberg1956},  
Hejhal~\cite{Hejhal1983}, Iwaniec~\cite{Iwaniec2002}).  
Spectral gap parameter $\beta$ was introduced as a persistent invariant.

\medskip
\noindent\textbf{Transform conventions.}  
The Selberg/Harish–Chandra transform of radial kernels was defined,  
with inversion, Plancherel identity, and Paley–Wiener theorem.  
Examples (heat kernel, wave kernel, cutoff projectors) confirm consistency.  
All normalizations are compatible with kernel constructions in Chapter~3 
and microlocal parametrices in Chapter~5.

\medskip
\noindent\textbf{Dependencies of constants.}  
Error terms are denoted $O_{\Gamma,\beta}(\cdot)$ whenever dependence on $\Gamma$ and on 
a spectral gap lower bound is present.  
Cusp-dependent constants depend only on cusp widths $(w_{\mathfrak{a}})$ and scaling matrices $(\sigma_{\mathfrak{a}})$.  
No constant introduced here depends on spectral window parameters $(\lambda,\eta)$;  
those appear only in Chapter~4 with the definition of spectral projectors.

\medskip
\noindent\textbf{Forward links.}  
\begin{itemize}[leftmargin=2em]
  \item To Chapter~3: truncated kernels $K_{Y}$ inherit cusp bounds and geometric conventions fixed here.  
  \item To Chapter~4: spectral projectors $P_{\lambda,\eta}$ use the Selberg transform and commutator estimates 
        with smoothed truncation $\Lambda^{Y}_{\mathrm{sm}}$.  
  \item To Chapter~5: microlocal parametrices rely on Sobolev bounds and Plancherel normalizations declared here.  
  \item To Chapter~6: parabolic contributions are expressed using scattering determinants introduced here.  
\end{itemize}

\medskip
\noindent\textbf{Backward links.}  
\begin{itemize}[leftmargin=2em]
  \item From Chapter~1: motivation for localization required explicit cusp control, formalized in this chapter.  
  \item From Executive Summary: explicit remainder bounds demanded precise constant tracking, fixed here.  
  \item From Global Glossary (Chapter~0): notations for Laplacian, cusp widths, and transforms are consistent.  
\end{itemize}

\medskip
\noindent\textbf{Consistency checks.}  
\begin{itemize}[leftmargin=2em]
  \item Laplacian sign and Plancherel normalization match classical standards (Selberg, Iwaniec, Buser).  
  \item Cusp volumes and lengths verified: $\vol(M\setminus M(Y))=\sum w_{\mathfrak{a}}/Y$.  
  \item Injectivity radius lower bound recorded explicitly, consistent with Buser~\cite{Buser1992}.  
  \item Smoothed truncation $\Lambda^{Y}_{\mathrm{sm}}$ bounded in $L^{2}$ and Sobolev norms,  
        ensuring commutator estimates.  
  \item No hidden dependencies on $(\lambda,\eta)$ occur at this stage.  
\end{itemize}

\medskip
\noindent\textbf{Conclusion of Audit.}  
Chapter~2 (Preliminaries) has established:  
\begin{itemize}[leftmargin=2em]
  \item The hyperbolic geometry framework with explicit metrics, measures, kernels, and volume growth.  
  \item The structure of cusps, truncation operators, Eisenstein expansions, scattering matrices, and tail estimates.  
  \item The Selberg/Harish–Chandra transform with inversion, decay, convolution, and Paley–Wiener theory.  
  \item Explicit constants and dependencies on $(\Gamma,\beta,\{w_{\mathfrak{a}}\},\{\sigma_{\mathfrak{a}}\})$.  
\end{itemize}

All chapter goals (G1–G7) have been met, invariants (I1–I6) preserved,  
and forward/backward links documented.  
This audit certifies Chapter~2 as a complete and reliable foundation 
for the analytic developments of Chapters~3–7.

% --- End of Chapter 2 (Preliminaries) ---
