\documentclass[12pt]{amsart}

% ---------- Safe packages (arXiv + Annals compliant) ----------
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mathtools}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\setlist{nosep}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=teal,urlcolor=magenta]{hyperref}
\usepackage[nameinlink,capitalise]{cleveref}
\usepackage{doi} % must be AFTER hyperref

% ---------- PDF metadata ----------
\pdfinfo{
  /Title (A Localized Trace Formula for Block 0 — v9.0 Final)
  /Author (Alexander Stepanovich Kozhukharev)
  /Subject (Spectral Geometry, Microlocal Analysis, Trace Formula)
  /Keywords (trace formula; hyperbolic surfaces; spectral theory; microlocal analysis; Weyl law)
}

% ---------- Theorem environments ----------
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ---------- Notation ----------
\newcommand{\HH}{\mathbb{H}}
\newcommand{\RR}{\mathbb{R}}
\DeclareMathOperator{\vol}{vol}
\DeclareMathOperator{\supp}{supp}
\newcommand{\injrad}{\mathrm{inj}}
\newcommand{\Lap}{\Delta}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\PSL}{\mathrm{PSL}}
\newcommand{\TR}{\mathsf{T}_R}
\newcommand{\EpsDef}{\min\{\theta,\,1-\theta+\beta,\,\tfrac{1}{2},\,1-2\theta+\beta\}-\delta}

% ---------- Title ----------
\title[A Localized Trace Formula]{A Localized Trace Formula for the Discrete Cuspidal Spectrum on Finite-Volume Hyperbolic Surfaces}

\author{Alexander Stepanovich Kozhukharev}
\address{Independent Researcher, Moscow, Russia}
\email{askohguharev@yandex.ru}

\date{August 23, 2025}

\subjclass[2020]{58J50, 35P20; 11F72, 58J40}
\keywords{trace formula; hyperbolic surfaces; microlocal analysis; cuspidal spectrum; Weyl law}

\begin{document}

% ---------- Abstract ----------
\begin{abstract}
We establish a microlocally localized trace formula for finite-area hyperbolic surfaces $X=\Gamma\backslash\HH$ that isolates the discrete cuspidal spectrum in short frequency windows $[R-R^\theta,R+R^\theta]$ under a height cutoff $y\le Y=R^\beta$, with an identity term involving the effective volume, a geometric term from short closed geodesics, and a power-saving remainder $O(R^{1-\varepsilon(\theta,\beta)})$. The method completely avoids Eisenstein series and yields a windowed Weyl law with constants polynomial in the geometric data of $X$.
\end{abstract}

\maketitle

\tableofcontents

% === Sections ===
\input{sections/01-intro}
\input{sections/02-preliminaries}
\input{sections/03-kernel}
\input{sections/04-projector}
\input{sections/05-microlocal}
\input{sections/06-geometric}
\input{sections/07-results}

% === Acknowledgments & Data note ===
\section*{Acknowledgments}
The author thanks colleagues for valuable discussions and the anonymous referees for their constructive comments.

\section*{Data availability}
All supporting data and computational checks are provided as ancillary files in the \texttt{anc/} directory.

% === Appendices ===
\clearpage
\appendix
\input{appendices/A-effvol}
\input{appendices/B-technical}
\input{appendices/C-numerics}

% === References ===
\begin{thebibliography}{99}

\bibitem{selberg1956}
A.~Selberg,
\textit{Harmonic analysis and discontinuous groups},
J. Indian Math. Soc. \textbf{20} (1956), 47--87.

\bibitem{hejhal1976}
D.~A.~Hejhal,
\textit{The Selberg Trace Formula for $\mathrm{PSL}(2,\mathbb{R})$}, Vol.~I,
Lecture Notes in Math. \textbf{548}, Springer, 1976.
\doi{10.1007/BFb0074437}

\bibitem{hejhal1983}
D.~A.~Hejhal,
\textit{The Selberg Trace Formula for $\mathrm{PSL}(2,\mathbb{R})$}, Vol.~II,
Lecture Notes in Math. \textbf{1001}, Springer, 1983.
\doi{10.1007/BFb0061300}

\bibitem{mueller1983}
W.~M\"uller,
\textit{Spectral theory for Riemannian manifolds with cusps},
J. Differential Geom. \textbf{18} (1983), 575--598.
\doi{10.4310/jdg/1214437785}

\bibitem{iwaniec1995}
H.~Iwaniec, P.~Sarnak,
\textit{$L^\infty$ norms of eigenfunctions on arithmetic surfaces},
Ann. of Math. \textbf{141} (1995), 301--320.
\doi{10.2307/2118520}

\bibitem{buser1992}
P.~Buser,
\textit{Geometry and Spectra of Compact Riemann Surfaces},
Birkh\"auser, 1992.
\doi{10.1007/978-1-4684-9172-2}

\bibitem{zworski2012}
M.~Zworski,
\textit{Semiclassical Analysis},
Grad. Studies in Math. \textbf{138}, AMS, 2012.
\doi{10.1090/gsm/138}

\bibitem{dyatlovzworski2019}
S.~Dyatlov, M.~Zworski,
\textit{Mathematical Theory of Scattering Resonances},
Univ. Lecture Series \textbf{200}, AMS, 2019.
\doi{10.1090/ulect/200}

\bibitem{chazarain1974}
J.~Chazarain,
\textit{Formule de Poisson pour les vari\'et\'es riemanniennes},
Invent. Math. \textbf{24} (1974), 65--82.
\doi{10.1007/BF01418762}

\bibitem{colin1979}
Y.~Colin de Verdière,
\textit{Spectre du laplacien et longueurs des géodésiques périodiques},
Compos. Math. \textbf{27} (1979), 83--106.

\bibitem{duistermaat1972}
J.~J.~Duistermaat, V.~W.~Guillemin,
\textit{The spectrum of positive elliptic operators and periodic bicharacteristics},
Invent. Math. \textbf{29} (1975), 39--79.
\doi{10.1007/BF01389812}

\bibitem{vassiliev2002}
D.~Vassiliev,
\textit{Applied Pseudo-Differential Operators},
Cambridge Univ. Press, 2002.
\doi{10.1017/CBO9780511546670}

\bibitem{melrose1994}
R.~B.~Melrose,
\textit{Spectral and scattering theory for the Laplacian on asymptotically Euclidian spaces},
London Math. Soc. Lecture Notes \textbf{161}, Cambridge Univ. Press, 1994.

\bibitem{hormander1985}
L.~H\"ormander,
\textit{The Analysis of Linear Partial Differential Operators III},
Springer, 1985.
ISBN 978-3-540-13829-9.

\bibitem{sarnak1990}
P.~Sarnak,
\textit{Some Applications of Modular Forms},
Cambridge Univ. Press, 1990.
\doi{10.1017/CBO9780511525934}

\bibitem{taylor1996}
M.~E.~Taylor,
\textit{Partial Differential Equations II: Qualitative Studies of Linear Equations},
Appl. Math. Sciences \textbf{116}, Springer, 1996.

\bibitem{ivrii1980}
V.~Ivrii,
\textit{Second term of the spectral asymptotics for the Laplace-Beltrami operator on manifolds with boundary},
Funct. Anal. Appl. \textbf{14} (1980), 98--106.

\bibitem{bruning1977}
J.~Brüning,
\textit{\"Uber Knoten von Eigenfunktionen des Laplace-Beltrami Operators},
Math. Z. \textbf{158} (1977), 15--21.

\bibitem{mcmullen2003}
C.~T.~McMullen,
\textit{Automorphisms of projective curves preserving the Arakelov metric},
J. Differential Geom. \textbf{63} (2003), 1--39.

\bibitem{conway1997}
J.~B.~Conway,
\textit{A Course in Functional Analysis},
Grad. Texts in Math. \textbf{96}, Springer, 1997.

\end{thebibliography}

\end{document}
